---
title: "Filtered Summary"
author: "Emily O'Grady"
date: "2024-05-23"
output: html_document
---

```{r setup}
library(librarian)
shelf(tidyverse, ggplot2, ocedata)
```
This document contains a visual summary of data gathered for the AZMP climatology (2024), after
being filtered based on QC methods outlined by Gordana Lazin (2014).

Quality Control included filtering values based on methods, proximity to coastline, existing
quality control flags from BioChem, and profile envelope limits (Test 2.4 of IML QC).

The scope of this data will be visualized and compared to the scope of data used by Lazin et al in the last climatology (2014).


```{r}

data_fns <- list.files('data/filtered_data/', pattern = '.csv', full.names = TRUE)

data14 <- data.frame(
  variable = c("Nitrate", "Silicate", "Phosphate", "Chlorophyll A"),
  total_records = c(71698, 71499, 74514, 58424))

for (i in 1:length(data_fns)) {
  data <- read.csv(data_fns[i])
  
  param_name <- unique(data$PARAMETER_NAME)
  cat(param_name, '\n')
  
  if (param_name %in% data14$variable){
    cat("Total records used in 2014 climatology: ")
    cat(data14$total_records[data14$variable == param_name], '\n')
  } else {
    cat("Variable not present in 2014 climatology. \n")
  }
  cat("Total records available: ")
  cat(nrow(data), '\n')
  
  
  cat('\n \n')
  
}

```





# Profile plots

Visualize profile for all climatology data.

```{r}
data_fns <- list.files('data/filtered_data/', pattern = '.csv', full.names = TRUE)


for (i in 1:length(data_fns)) {
  data <- read.csv(data_fns[i])
  param_name <- unique(data$PARAMETER_NAME)
  
  p_prof <- ggplot(data) +
    geom_point(aes(x = DATA_VALUE, y = HEADER_START_DEPTH))+
    scale_y_reverse()+
    labs(x = 'Data Value', y = 'Depth') +
    scale_x_continuous(position = 'top')+
    ggtitle(paste(param_name, 'profile of all data for climatology'))
    
  print(p_prof)
}


```


# Map plots

```{r}
boxdat <- read.csv("Gordana/2014_boxes.csv")
boxdat <- boxdat %>%
  group_by(box)

data(coastlineWorldFine)
clwf <- as.data.frame(coastlineWorldFine@data)


data_fns <- list.files('data/filtered_data/', pattern = '.csv', full.names = TRUE)


for (i in 1:length(data_fns)) {
  data <- read.csv(data_fns[i])
  param_name <- unique(data$PARAMETER_NAME)
  
  pp <- ggplot()+
    geom_polygon(data = boxdat,
                 aes(x = longitude, y = latitude, fill = as.factor(box)),
                 colour = 'black')+
    geom_polygon(data = clwf,
                 aes(x = longitude, y = latitude),
                 colour = 'grey')+
    geom_point(data = data,
               aes(x = EVENT_MIN_LON, y = EVENT_MIN_LAT),
               shape = 21,
               colour ='black',
               fill = 'darkgrey',
               alpha = 0.2) +
    scale_x_continuous(limits = c(-71, -48),
                       expand = c(0,0))+
    scale_y_continuous(limits = c(37, 48),
                       expand = c(0,0)) +
    labs(fill = 'Box')+
    theme_classic()+
    ggtitle(paste('Spatial distribution of', param_name))
  
  print(pp)
}

```





