---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(oce)
library(ocedata)
library(tidyverse)
library(sf)
```

# AZMP_climatology

<!-- badges: start -->
<!-- badges: end -->

The goal of AZMP_climatology is to generate a monthly climatology product
for the Scotian Shelf. This product will be used to update the QC methodology 
used at Bedford Institute of Oceanography to quality control bottle data from
our long term monitoring programs. 

The current QC system compares bottle data to a monthly climatology which was 
generated in 2014 by Gordana Lazin, Andrew Cogswell, Catherine Johnson, 
Shelley Bond and Jeff Jackson (unpublished). This climatology makes some small 
improvements in data quality and parameter coverage. It also updates the dataset
to a more relevant time period. 

# 2014 comparison

There were multiple changes implemented in this updated version of the Scotian
Shelf climatology, although most of the methods remained the same as they were
in 2014. The main changes were:

- A reduced temporal range
        Data was only extracted from 1999-2023, as opposed to 1950-2013 in the 2014 version

- Additional quality control
        As recommended by the original authors in 2014, additional QC procedures
        were implemented before climatology calculations were performed, including 
        stage 2 testing from the IML QC protocol (range testing, comparitive 
        measurements, constant profile, replicate variation, and others)

- Improved data coverage
        More parameters are included in this updated version. Phaeo, Ammonia, 
        Nitrite, oxygen, were added to the list of parameters included in the
        climatology.
        


# Data summary

The data used in this project was extracted from BioChem, DFO's national database. 
Data for relevant parameters was extracted for the Scotian Shelf region. The data
was then put through a series of quality control steps to verify quality and remove
outliers and questionable data. Data was then summarized in 9 geographic boxes, 
over 4 depth intervals, for each month. 


```{r load-data, echo=FALSE}
# load data24

varfiles <- list.files('data/filtered_data', pattern = 'final', full.names = TRUE)
for (i in 1:length(varfiles)) {
  # load all files and collate them into one dataframe
  if (i == 1) {
    data24 <- read.csv(varfiles[i])
  } else {
    data24 <- rbind(data24, read.csv(varfiles[i]))
  }
  
}


data(coastlineWorldFine)
clwf <- as.data.frame(coastlineWorldFine@data)

# read in boxes
boxdat <- read.csv("Gordana/2014_boxes.csv")
boxdat <- boxdat %>%
  group_by(box)

box_poly <- st_as_sf(boxdat, coords = c("longitude", "latitude"), crs = 4326)
# Create polygon
df_poly <- box_poly %>%
  group_by(box) %>%
  summarise(geometry = st_combine(geometry), do_union = FALSE) %>%
  st_cast("POLYGON")

```


```{r spatial-distirbution, echo = FALSE, warning=FALSE, message = FALSE}

# plot spatial distribution of dataset



# plot spatial distribution of data24 with coastline and climatology boxes

ggplot()+
  geom_path(data = clwf, aes(x = longitude, y = latitude), colour = 'black', fill = 'grey')+
  geom_sf(data = df_poly, colour = 'black', alpha = 0.5)+
  geom_point(data = data24, aes(x = HEADER_START_LON, y = HEADER_START_LAT), colour = 'darkblue', alpha = 0.1)+
  ggtitle('Spatial distribution of data24')+
  theme_classic()+
  theme(legend.position = 'none') +
  coord_sf(xlim = c(-71, -48), ylim = c(37, 48))



```

Figure 1. Spatial distribution of all data included in the climatology. Scotian 
Shelf climatology boxes are shown in grey.


```{r temporal-distribution, echo = FALSE, warning=FALSE, message = FALSE}

# plot temporal distribution of entire dataset

yearMonth <- data24 %>%
    mutate(., year = format(as.Date(data24$EVENT_START, '%d-%b-%y'), '%Y')) %>%
    mutate(., month = format(as.Date(data24$EVENT_START, '%d-%b-%y'), '%b')) %>%
    group_by(year, month) %>%
    summarise(N = n())
  
ggplot(yearMonth, aes(year, month)) +
    geom_tile(aes(fill = N), colour = "white") +
    scale_fill_gradient(low = "blue", high = "red") +
    labs(x = "Year", y = "Month", fill = "Number of \n Data Points") +
    theme_minimal()+
    ggtitle(paste("Temporal coverage"), subtitle = paste("Total points: ", sum(yearMonth$N)))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Figure 2. Temporal distribution of climatology data. Each tile represents the 
number of data points collected in a given month. 


## Reports and scripts available

The workflow followed to produce this climatology followed:
  - initial_summary.Rmd which gave an intial scope of the extracted data
  - qc.rmd where data was quality controlled and excluded based on testing
  - filtered_summary.rmd where QC'd data was summarized and visualized
  - report.rmd summarizes the QC'd data in a report format
  - visualize_final_ranges_2024.rmd where final data was summarized into climatology statistics
  - final_climatology_calculation.Rmd where climatology was calculated using 2014 data to fill data gaps
  
Other scripts were used to generate additional plots and investigate other data aspects.

Some work was done to investigate the inclusion of CTD chlorophyll A in the climatology
  (process_ctd_Data.R, 01,_readData.R, 02_inferChlorophyllInstrument.R)
This avenue was not further pursued due to high effort, and low potential for improvement in the climatology.

Some work was also done to investigate the inclusion of TS data for an updated 
physical climatology. 
  (TS_climatology_CASTS.R, ts_climatology.R)
This could be further pursued in the future but was beyond the capacity of this project.



## Products available

The 'data' folder contains multiple stages of data for each parameter. 
  - initial data extraction
  - QC'd data
  - filtered data
  - climatology data
  
TS climatology data (CASTS) is also available, and other data extracted from NOAA for temperature and salinity. 

The 'Gordana' folder, includes documentation and data products from the 2014 climatology. 


## Update Log



